### 2008-08-14 doc/a-rex/arex_tech_doc.tex
More changes in a-rex document

\textbf{You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
instalaltion and configuration refer to other documents avaialble
at \url{http://www.nordugrid.org/papers.html}. }

@@

A job is a set of input files (which may or may not include executables),
a main executable and a set of output files. The process of gathering
input files, executing a job, and transferring/storing output files
is called a \emph{session}.

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files  in the SD as well. The A-Rex does not guarantee
the availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the A-Rex.
It is  accessible by the user from outside through HTTP(S)  
protocol. Any file created outside the SD is not controlled by the
A-Rex. Any exchange of data between client and A-Rex (including
also program modules) is done through    HTTP(S) protocol
   . A URL for accessing input/output
files is       obtained through
the WS Local Information Description Interface (LIDI) of A-Rex.

Each job gets an identifier (\textit{jobid}). This is a handle which
identifies the job in the A-Rex and the Information  Interface.

  Jobs are initiated and controlled through   WS interface.
Complete job descriptions   (JD)   are passed to the
A-Rex  through       WS in JSDL \cite{jsdl}
coded description     . Input data files
and job executables are transferred separately through the same interface,
as described in the next section.

@@

One of the most important tasks   of the A-Rex is to
take care of processing input and output data (files) of the job.
Input files are gathered in the SD. There are two ways to put a file
into the SD:

\begin{itemize}
\item Download is initiated by the A-Rex --  This is the case
for files     defined in the JD (with name and source).
      The A-Rex alone is responsible to
make sure that all required files will be available in the SD.\\
The supported protocols for sources  at the moment are: GridFTP,
FTP, HTTP, HTTPS (HTTP over SSLv3) and HTTPg (HTTP over GSI). Also
some nonstandard sources are supported. Those are described below.
The A-Rex fully relies on HED framework \cite{hed} for data transfering
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-Rex.  Examples of
such files  are the main executable of the job,  
the job's input files, etc. These files can (and
should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-Rex  . Access to
AFS, NFS, FTP, HTTP and any other remote data transport during execution
of a job is not guaranteed (at least not by A-Rex).

@@

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service } (RLS)  --
The A-Rex takes care of these files. They have to be specified
in the JD. If the job fails during any stage of processing, no attept
is made to transfer those files to their final destination, unless
the  option \emph{preserve=yes} is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}

@@

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the dedicated Local
Information Description Interface of A-Rex  
 using the UI or any other suitable tool

     or through query method of WS interface.

Configuration can put limits on the amout of simultaneous jobs in
some states. If such a limit is reached, a job ready to enter into
the state in question will stay in it's current state waiting for
a free slot. This situation is presented by prepending the prefix
\textbf{PENDING:} to the current state name  \textbf{}
in the job's status mark.

@@

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to
a CE but is not processed yet. The A-Rex will analyze the JD and
move to the next stage. If the JD can not be processed the job will
be canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-Rex is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If
all files are successfully gathered the job moves to the next state.
If \textbf{any} file can't be downloaded or it takes the UI too long
to upload a file,  the job moves to \textbf{Finishing} state. It
is possible to put a limit on the number of simultaneous \textbf{Preparing}
jobs. If this limit is exceeded, jobs ready to enter the \textbf{Preparing}
state      will stay in the \textbf{Accepted}
state, but prefixed with the PENDING: mark. Exceptions are jobs which
have no files to be downloaded. These are processed out of limits.
\item \textbf{Submitting} -      The job is being
passed for execution to the \emph{Local Resource Management System}
(LRMS).         The correspoding backends
for many LRMSs are provided with the default installation.  
     If the local job submission is successful
the job moves to the \textbf{InLRMS} state. Otherwise it moves
to \textbf{Finishing}. It is possible to limit the agregate number
of jobs in \textbf{Submitting} and  \textbf{InLRMS} states.
\item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
The A-Rex takes no actions except waiting until the job finishes.
\item \textbf{Cancelling} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at  an Indexing Service. The user can download data files
from the SD by using the UI or  other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-Rex.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the
job is moved to the state \textbf{Deleted}. The 'deletion' time can
be obtained by querying the Informatrion Interface of the A-Rex
      . If a job was moved to \textbf{Finished}
because of failure, it may be restarted on request of a client. When
restarted, a job is moved to the state previous to the one in which
it failed and is assigned mark PENDING. This is needed in order to
not break the configuration limits. Exception is a job failed in \textbf{InLRMS}
state and lacking input files specified in JD. Such a job is treated
like failed in \textbf{Preparing} state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In  case of a failure special processing is applied to output
files. All specified output files are treated as \textbf{downloadable
by the user}. No files will be moved to their destination SE.

@@

In full installation the A-Rex and it's components support the
following data transfer protocols and corresponding URLs: \emph{ftp,
gsiftp, http, httpg, https, se, rc} and \emph{rls.} For more information
please see {}``The Hosting Environment of the Advanced Resource Connector
middleware      
  '' document \cite{hed}.

@@

\subsection{Internal Files of the A-Rex }

For each local UNIX user listed in the A-Rex configuration - including
a generic one which covers all local user identities - a \textit{control
directory} exists. In this directory the A-Rex stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. To make it easier to recover in  case
of failure, the A-Rex stores most information in files rather than
in memory. All files belonging to the same job have names starting
with \textbf{job.ID.,} where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. This is a plain
text file containing a single word   representing the current
state of the job. Possible values are :

\begin{itemize}
\item ACCEPTED
\item PREPARING
\item SUBMITTING
\item INLRMS
\item FINISHING
\item FINISHED
\item CANCELING
\item DELETED
\end{itemize}
\end{itemize}
See section \ref{Section: Job Flow} for a description of the various
states. Additionally each value can be prepended  the prefix
{}``PENDING:'' (like PENDING:ACCEPTED, see section \ref{Section: Job Flow}).
This is used to show that a job  is \emph{ready} to be moved
to the next state but it has to stay in it's current state \emph{only}
because otherwise some limits set in the configuration would be
exceeded.

\begin{itemize}
\item \textit{job.ID.description} - contains the  description of the
job (JD).
\item \textit{job.ID.local} - information about the job used by the A-Rex.
It consists of lines of format \textit{{}``name = value''} . Not
all of them are always available. The following names are defined:

\begin{itemize}
\item \textit{subject} - user certificate's subject, also known as the distinguished
name (DN)
\item \textit{starttime} -  GMT time when the job was accepted represented
in the Generalized Time format of LDAP 
\item \textit{lifetime} - time period to  preserve the SD after the
job has finished in seconds
\item \textit{cleanuptime} -  GMT time when the job should be removed
from the cluster and it's SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
the job specified status changes
\item \textit{processtime} -  GMT time when to start processing the job
in Generalized Time format
\item \textit{exectime} -  GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} -  GMT time when the credentials delegated
to the job expire in Generalized Time format
\item \textit{rerun} - number of retries left to rerun the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of the LRMS backend     to be used
for local submission
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only after the job
 has reached state \textbf{InLRMS})
\item \textit{args} - executable name followed by a list of command-line
arguments   
\item \textit{downloads} - number of files to download into the SD before
execution
\item \textit{uploads} - number of files to upload from the SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about the job when accessed through GridFTP interface
\item \textit{clientname} - name (as provided by the user interface) and
ip address:port of the submitting client machine

\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD  
\item \textit{failedstate} - state in which job failed (available only
if it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The GM will also send job records to this service in addition to the
default logger service configured in the configuration  
        (This feature does
not work in current implementation).
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

\textit{url} - ordinary URL for gsiftp, ftp, http, https or httpg
protocols with the addition of '\textbf{replica catalog url}' (RC
URL) and '\textbf{replica location service url}' (RLS URL).\\
Each URL can contain additional options.

@@

\item \textit{job.ID.proxy} -delegated X509 credentials .
\item \textit{job.ID.proxy.tmp} - temporary X509 credentials  with
different UNIX ownership used by processes run with effective \emph{user
id} different form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the A-Rex. Their presence in the
\textit{control directory} can not be guaranteed and can change depending
on changes in the A-Rex code.

\subsection{ }

    \textit{ }  \textit{}

 \textit{}

 \textit{}

 \textit{}

@@

A-R introduces a new operation in addition to those provided by
BES. It does that by defining its own port-type  with the single
operation \emph{ChangeActivityStatus} (see Appendix D). This operation
provides a way to request simple transfers between states of jobs
  and corresponding actions.

@@

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-Rex defines a set of sub-states of
activity processing in addition to those defined by the BES, as listed
in Table \ref{tab:Job-states-definitions}.

@@

\begin{tabular}{|c|c|>{\raggedright}p{10cm}|}
\hline 
Applicable BES State&
ARC Sub-state&
\multicolumn{1}{c|}{Description}\tabularnewline
\hline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{2}{2cm}{Pending}&
Accepting&
Job is in the process of being submitted\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Accepted&
Job was submitted\tabularnewline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{7}{2cm}{Running}&
Preparing&
Stage-in process is going on\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Prepared&
Stage-in process has finished\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Submitting&
Communication with local batch system is in process\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Executing&
Job is being executed in local batch system\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Killing&
Communication with local batch system to terminate execution is in
process\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Executed&
Job execution in local batch system has finished\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Finishing&
Stage-out process is going on\tabularnewline
\hline 
Cancelled&
Failed&
There was a failure during execution\tabularnewline
\cline{1-1} 
Failed&
\multicolumn{1}{c|}{}&
\multicolumn{1}{l|}{}\tabularnewline
\hline 
All&
Pending&
Job is prevented from going to the next state due to some internal
limits; this sub-state appears in parallel with other sub-states\tabularnewline
\hline 
All&
Held&
Job processing is suspended on client request; this sub-state appears
in parallel with other sub-states\tabularnewline
\hline
\end{tabular}
\end{table}

@@

\begin{itemize}
\item \emph{UnsupportedFault}: Indicates that service does not support this
operation despite supporting the port-type.
\item \emph{ProcessingFault}: Internal problems during generation of the
token.
\end{itemize}
\end{itemize}
\end{itemize}
Additionally, A-Rex Web Service Interface allows delegation to be
performed as part of the \emph{CreateActivity} operation of the BES-Factory
port-type. For this it accepts the element \emph{DelegatedCredentials}
inside the \emph{CreateActivity} element. The \emph{Id} element of
\emph{DelegatedCredentials} must contain an identifier obtained in
response to the previous \emph{DelegateCredentialsInit} operation.

@@

\begin{lyxlist}{00.00.0000}
\item [{\emph{BESFactory}}] - collection of BES Factory attributes as defined
in the BES specifications.
\item [{\emph{Glue2Resource}}] \emph{-} description of a computation resource
that uses Glue2 schema. This one is going to be specified after the
Glue2 XML bindings will be available.
\item [{\emph{Activities}}] \emph{-} list of maintained activities. Each
activity contains an identifier (\emph{ActivityIdentifier}), BES (\emph{ActivityDocument})
and Glue2 (\emph{Glue2Job}) descriptions of the activity.
\end{lyxlist}
All information can be accessed either through requests on particular
resources or through XPath queries using WS-Resource Properties operations.

@@

Before giving access to an already cached file, the GM contacts the
original file source to check if the user     has
suficient access rights to the file. Not all protocols  
 make this check possible.

Also, file creation or validity times are checked to make sure the
cached file is fresh . If it is impossible to obtain creation
and invalidation times  , the file is invalidated 24 hours
after it was downloaded.

@@

\begin{itemize}
\item \textit{grid-manager} - The main module. It is responsible for processing
 jobs, moving them through states and running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at  an Indexding
Service (like RLS) as needed. It processes and updates the \textit{job.ID.output}
file.
\item \emph{cache-register} - Utility to register cached data into an Indexing
Service    . It reads and modifies cache informational
files \emph{old} and \emph{new} (as described in section \ref{sec:cache}).
Configuration is read directly from the GM's configuration file (see
section \ref{SubSection: ConfigFile}). It is run by the GM every
5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
the frontend  . It puts  collected information into the \emph{job.ID.diag}
file.
\item \emph{gm-kick} - Sends a signal to the GM though a FIFO file to wake
it up. It's used to increase responsiveness of GM.
\end{itemize}
The following modules are always run under the Unix account to which
a grid user is mapped.

@@

\begin{itemize}
\item \textit{session root directory} -  Tthis is the directory under
which a user's SDs are created. It's location is configurable
per UNIX user. Several (or even all) users may share the same session
root directory.        
    .\\
There are 2 processes which need to have  permissions to create new
files and directories in it  : GM and GFS. \\
If any of these processes  are run under a dedicated user
account, that account needs full permissions in the \textit{session
root directory}.\\
If these processes are run under the \emph{root} account, make sure
\textit{session root directory}  resides on a filesystem which
does not limit the capabilities of the \emph{root} user  (as does
for example NFS with \emph{root\_squash} option).\\
If there is a need to run processes under the \emph{root} account
(to be able to run jobs in LRMS under different users' accounts, for
example) but there is no way to provide a suitable \textit{session
root directory,} use the \emph{norootpower} command in GM's configuration
  file. In that case GM and GFS will use the identity of the
local user to which a Grid identity is mapped to access the \textit{session
root directory}. Hence those users will need full access there.\\
The GM creates SDs with proper ownership and permissions for the local
identity used to run a job. Some filesystems require users to have
\emph{execute} permission on the \textit{session root directory}
      in order to access any file or subdirectory
there.\\
         In order
for jobs to access their input files, session root directories should
be shared across cluster nodes. Otherwise,    LRMS-specific
methods must be used to transfer files to execution nodes. For more
information see section \ref{sub:LRMS}.
\item \textit{control directory} - In this directory  GM and GFS store
an information about  accepted jobs. Both    processes
must have full permissions there.\\
A subdirectory called \textit{log} is created there. It is used
to accumulate information about started and finished jobs. This information
is periodically sent to the desired \emph{logger service}(s). For
each job start and stop event, and for each logger service where that
event must be sent, a separate file is written. Once an event is sent,
the corresponding file is deleted.
\end{itemize}

@@

\subsection{ Running as non-root}

>>>>

\textbf{You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
instalaltion and configuration of whole system refer to other documents
avaialble at \url{http://www.nordugrid.org/papers.html}. }

@@

A job is a set of input files (which may include executables), a main
executable and a set of output files. The process of gathering input
files, executing a job, and transferring/storing output files is called
a \emph{session}.

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files in the SD as well. The A-Rex does not guarantee the
availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the A-Rex. It
is accessible by the user from outside through HTTP(S) protocol. Any
file created outside the SD is not controlled by the A-Rex. Any exchange
of data between client and A-Rex (including also program modules)
is done through HTTP(S) protocol. A URL for accessing input/output
files is  obtained through the WS Local Information Description Interface
(LIDI) of A-Rex.

Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
\cite{jsdl} XML document - which identifies the job in the A-Rex
and the Information Interface.

Jobs are initiated and controlled through WS interface. Complete job
descriptions (JD)are passed to the A-Rex through WS in JSDL\cite{jsdl}
coded description. Input data files and job executables are transferred
separately through the same interface, as described in the next section.

@@

One of the most important tasks of the A-Rex is to take care of processing
input and output data (files) of the job. Input files are gathered
in the SD or associated cache area. There are two ways to put a file
into the SD:

\begin{itemize}
\item Download is initiated by the A-Rex -- This is the case for files defined
in the JD (with name and source). The A-Rex alone is responsible to
make sure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-Rex fully relies on HED framework \cite{hed} for data transfering
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-Rex. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-Rex. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-Rex).

@@

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service} (RLS) -- The A-Rex takes care
of these files. They have to be specified in the JD. If the job fails
during any stage of processing, no attept is made to transfer those
files to their final destination, unless the option \emph{preserve=yes}
is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}
All files not specified in the JD are deleted after job execution
finished. If job execution fails for any reason (if exit code of main
executable is not 0) all files from furst group are transfered to
second one.

@@

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the dedicated Local
Information Description Interface of A-Rex using the UI or any other
suitable tool or through query method of WS interface.

Configuration can put limits on the amout of simultaneous jobs in
some states. If such a limit is reached, a job ready to enter into
the state in question will stay in it's current state waiting for
a free slot. This situation is presented by additional state mark
\textbf{PENDING} to the current state name in the job's status desription.

@@

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to a
CE but is not processed yet. The A-Rex will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-Rex is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. If this
limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submiting} - The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The correspoding backends
for many LRMSs are provided with the default installation. If the
local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the agregate number of jobs in \textbf{Submiting} and \textbf{Executing}
states.
\item \textbf{Executing} - The job is queued or being executed in the LRMS.
The A-Rex takes no actions except waiting until the job finishes.
\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-Rex.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Informatrion Interface of the A-Rex. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

@@

In full installation the A-Rex and it's components support the following
data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
http, https, lfc} and \emph{rls.} For more information please see
{}``The Hosting Environment of the Advanced Resource Connector middleware''
document \cite{hed}.

@@

\subsection{Internal Files of the A-Rex}

For each local UNIX user listed in the A-Rex configuration - including
a generic one which covers all local user identities - a \textit{control
directory} exists. In this directory the A-Rex stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. In most common configuration case A-Rex
serves all users defined by Operating System and stores their control
file in same directory. To make it easier to recover in case of failure,
the A-Rex stores most information in files rather than in memory.
All files belonging to the same job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. This is a plain
text file containing a single word representing the internal name
of current state of the job. Possible values and corrsponding external
job states are :

\begin{itemize}
\item ACCEPTED/\textbf{Accepted}
\item PREPARING/\textbf{Preparing}
\item SUBMITTING/\textbf{Submiting}
\item INLRMS/\textbf{Executing}
\item FINISHING/\textbf{Finishing}
\item FINISHED/\textbf{Finished}
\item CANCELING/\textbf{Killing}
\item DELETED/\textbf{Deleted}
\end{itemize}
\end{itemize}
See section \ref{Section: Job Flow} for a description of the various
states. Additionally each value can be prepended the prefix {}``PENDING:''
(like PENDING:ACCEPTED, see section \ref{Section: Job Flow}). This
is used to show that a job is \emph{ready} to be moved to the next
state but it has to stay in it's current state \emph{only} because
otherwise some limits set in the configuration would be exceeded.

\begin{itemize}
\item \textit{job.ID.description} - contains the description of the job
(JD).
\item \textit{job.ID.local} - information about the job used by the A-Rex.
It consists of lines of format \textit{{}``name = value''}. Not
all of them are always available. The following names are defined:

\begin{itemize}
\item \textit{subject} - user certificate's subject, also known as the distinguished
name (DN)
\item \textit{starttime} - GMT time when the job was accepted represented
in the Generalized Time format of LDAP 
\item \textit{lifetime} - time period to preserve the SD after the job has
finished in seconds
\item \textit{cleanuptime} - GMT time when the job should be removed from
the cluster and it's SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
the job specified status changes
\item \textit{processtime} - GMT time when to start processing the job in
Generalized Time format
\item \textit{exectime} - GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} - GMT time when the credentials delegated to the
job expire in Generalized Time format
\item \textit{rerun} - number of retries left to rerun the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of the LRMS backend to be used for local submission
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only after the job has
reached state \textbf{InLRMS})
\item \textit{args} - executable name followed by a list of command-line
arguments
\item \textit{downloads} - number of files to download into the SD before
execution
\item \textit{uploads} - number of files to upload from the SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about the job when accessed through GridFTP interface
\item \textit{clientname} - name (as provided by the user interface) and
ip address:port of the submitting client machine
\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD
\item \textit{failedstate} - state in which job failed (available only if
it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The GM will also send job records to this service in addition to the
default logger service configured in the configuration (This feature
does not work in current implementation).
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

\textit{url} - ordinary URL for gsiftp, ftp, http, https protocols
with the addition of '\textbf{replica location service url}' (RLS
URL).\\
Each URL can contain additional options.

@@

\item \textit{job.ID.proxy} -delegated X509 credentials.
\item \textit{job.ID.proxy.tmp} - temporary X509 credentials with different
UNIX ownership used by processes run with effective \emph{user id}
different form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the A-Rex. Their presence in the \textit{control
directory} can not be guaranteed and can change depending on changes
in the A-Rex code.

@@

A-Rex introduces a new operation in addition to those provided by
BES. It does that by defining its own port-type with the single operation
\emph{ChangeActivityStatus} (see Appendix D). This operation provides
a way to request simple transfers between states of jobs and corresponding
actions.

@@

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-Rex defines a set of sub-states of activity
processing in addition to those defined by the BES, as listed in Table
\ref{tab:Job-states-definitions}. Their meaning is described in section
\ref{Section: Job Flow}.

@@

\begin{tabular}{|c|c|>{\raggedright}p{10cm}|}
\hline 
Applicable BES State&
ARC Sub-state&
\multicolumn{1}{c|}{Description}\tabularnewline
\hline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{2}{2cm}{Pending}&
Accepting&
Job is in the process of being submitted\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Accepted&
Job was submitted\tabularnewline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{7}{2cm}{Running}&
Preparing&
Stage-in process is going on\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Prepared&
Stage-in process has finished\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Submiting&
Communication with local batch system is in process\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Executing&
Job is being executed in local batch system\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Killing&
Communication with local batch system to terminate execution is in
process\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Executed&
Job execution in local batch system has finished\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Finishing&
Stage-out process is going on\tabularnewline
\hline 
Cancelled&
Failed&
There was a failure during execution\tabularnewline
\cline{1-1} 
Failed&
\multicolumn{1}{c|}{}&
\multicolumn{1}{l|}{}\tabularnewline
\hline 
All&
Pending&
Job is prevented from going to the next state due to some internal
limits; this sub-state appears in parallel with other sub-states\tabularnewline
\hline 
All&
Held&
Job processing is suspended on client request; this sub-state appears
in parallel with other sub-states. This state is reserved for future
and i snot implemented yet.\tabularnewline
\hline
\end{tabular}
\end{table}

@@

\begin{itemize}
\item \emph{UnsupportedFault}: Indicates that service does not support this
operation despite supporting the port-type.
\item \emph{ProcessingFault}: Internal problems during generation of the
token.
\end{itemize}
\end{itemize}
\end{itemize}
Additionally, A-Rex Web Service Interface allows delegation to be
performed as part of the \emph{CreateActivity} operation of the BES-Factory
port-type. For this it accepts the element \emph{DelegatedCredentials}
inside the \emph{CreateActivity} element. The \emph{Id} element of
\emph{DelegatedCredentials} must contain an identifier obtained in
response to the previous \emph{DelegateCredentialsInit} operation.
For more information about delegations and delegation interface refer
to \cite{wsrf-rp}.

@@

\begin{lyxlist}{00.00.0000}
\item [{\emph{nordugrid}}] - desription of computing resource that uses
NorudGrid LDAP schema {[}] converted to XML document.
\item [{\emph{Domains}}] \emph{-} description of a computation resource
that uses Glue2 schema.
\end{lyxlist}
All information can be accessed either through requests on particular
resources or through XPath queries using WS-Resource Properties operations.

@@

Before giving access to an already cached file, the GM contacts the
original file source to check if the user has suficient access rights
to the file. Not all protocols make this check possible. Also, file
creation or validity times are checked to make sure the cached file
is fresh. If it is impossible to obtain creation and invalidation
times, the file is invalidated 24 hours after it was downloaded.

@@

\begin{itemize}
\item \textit{grid-manager} - The main module. It is responsible for processing
jobs, moving them through states and running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at an Indexding Service
(like RLS) as needed. It processes and updates the \textit{job.ID.output}
file.
\item \emph{cache-register} - Utility to register cached data into an Indexing
Service. It reads and modifies cache informational files \emph{old}
and \emph{new} (as described in section \ref{sec:cache}). Configuration
is read directly from the GM's configuration file (see section \ref{SubSection: ConfigFile}).
It is run by the GM every 5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
the frontend. It puts collected information into the \emph{job.ID.diag}
file.
\item \emph{gm-kick} - Sends a signal to the GM though a FIFO file to wake
it up. It's used to increase responsiveness of GM.
\end{itemize}
The following modules are always run under the Unix account to which
a grid user is mapped.

@@

\begin{itemize}
\item \textit{session root directory} - Tthis is the directory under which
a user's SDs are created. It's location is configurable per UNIX user.
Several (or even all) users may share the same session root directory..\\
There are 2 processes which need to have permissions to create new
files and directories in it: GM and GFS. \\
If any of these processes are run under a dedicated user account,
that account needs full permissions in the \textit{session root directory}.\\
If these processes are run under the \emph{root} account, make sure
\textit{session root directory} resides on a filesystem which does
not limit the capabilities of the \emph{root} user (as does for example
NFS with \emph{root\_squash} option).\\
If there is a need to run processes under the \emph{root} account
(to be able to run jobs in LRMS under different users' accounts, for
example) but there is no way to provide a suitable \textit{session
root directory,} use the \emph{norootpower} command in GM's configuration
file. In that case GM and GFS will use the identity of the local user
to which a Grid identity is mapped to access the \textit{session root
directory}. Hence those users will need full access there.\\
The GM creates SDs with proper ownership and permissions for the local
identity used to run a job. Some filesystems require users to have
\emph{execute} permission on the \textit{session root directory} in
order to access any file or subdirectory there.\\
In order for jobs to access their input files, session root directories
should be shared across cluster nodes. Otherwise, LRMS-specific methods
must be used to transfer files to execution nodes. For more information
see section \ref{sub:LRMS}.
\item \textit{control directory} - In this directoryGM and GFS store an
information about accepted jobs. Both processes must have full permissions
there.\\
A subdirectory called \textit{log} is created there. It is used to
accumulate information about started and finished jobs. This information
is periodically sent to the desired \emph{logger service}(s). For
each job start and stop event, and for each logger service where that
event must be sent, a separate file is written. Once an event is sent,
the corresponding file is deleted.
\end{itemize}

@@

\subsection{Running as non-root}
